---
title: "NMDS"
author: "Daphnée Lamarche"
date: "27 avril 2016"
output: html_document
---
### Make ordination via phyloseq
```{r}

set.seed(1234)

nmdsk4.ord <- ordinate(data, method = "NMDS", distance = "bray", k=4)

```
###Make ordination via vegan using dm (not able to figure out how to use raw data ... directly.. read someplace that it needs to be in log transformed ??)
```{r}
BC_Dm <- phyloseq::distance(data, method = "bray")
BC_DM = distance(rel_data, method="bray")

## Default is k=2
BC2.nmds<-metaMDS(BC_DM, k=2)
BC3.nmds<-metaMDS(BC_DM, k=3)
BC4.nmds<-metaMDS(BC_DM, k=4)
BC5.nmds<-metaMDS(BC_DM, k=5)

```
### uses different k values to determine how many dimensions you will need to have an optimal stress value (under 0.2 and fist point where the diffence is huge)
```{r}
BC2.nmds
# stress = 0.2214219
BC3.nmds
# stress =  0.1595712 
BC4.nmds
# stress = 0.1267351
BC5.nmds
# stress = 0.1046352
# I will use BC3.nmds (k=3)
```
### Make stressplot, make sure your points are scatter close to the line, and that your point fit the linear line (model??)
```{r}
stressplot(BC3.nmds)
```
## Now look at the 3rd axis to be sure you're not missing anything
```{r}

p <- plot_ordination(rel_data, BC3.nmds, type = "samples", color = "Type", axes = c(1,3),title = "Bray-Curtis Baseline and healthy")

p + geom_point(size = 3.5) + theme(legend.title = element_text(size=20)) + theme(legend.text = element_text(size=18)) + plot(vec.sp, p.max=0.01, col='blue')

```
### Get the points (cooordinate value)

```{r}
BC3.nmds$point

write.table(BC3.nmds$point, "BC3.nmds.csv", sep = ",")

```

