GBS final stats analysis
========================================================
Jennifer C. Stearns,
`r as.character(format(Sys.Date(), format="%B %d, %Y"))`

```{r, warning=FALSE, message=FALSE}
setwd("/home/jstearns/Documents/Baby_mi/")
library(knitr)
opts_chunk$set(cache=FALSE)
library(vegan)
library(ggplot2)
library(lme4)
library(grid)
library(lmerTest)
library(effects)
library(reshape2)
```

Notes:
* use BIC method to choose a model (not done)
* for parity groups are: 0, 1 and 2 or more kids
* for probiotics all types are glommed together as Y 

From this [R code for ecology tutorial](http://userweb.eng.gla.ac.uk/umer.ijaz/bioinformatics/ecological.html)
```{r}
meta_table<-read.csv("sub_meta.csv", row.names = 1, na.strings="#NULL!")

meta_table$age = factor(meta_table$age,levels(meta_table$age)[c(3,1,4,2)])
meta_table$age <- factor(meta_table$age,ordered=TRUE)

hist(meta_table$actual_age, breaks=50,xlim=c(0,110), plot=TRUE,col="grey", xlab="age in days")
meta_table[,5:23] <- lapply(meta_table[,5:23] , factor)
# rescale aage to weeks
meta_table$actual_age <- meta_table$actual_age/7
hist(meta_table$actual_age, breaks=15, plot=TRUE,col="grey", xlab = "age in weeks")
meta_table$ipa_duration <- meta_table$ipa_duration/60
hist(meta_table$ipa_duration, plot=TRUE,col="grey", xlab = "ipa duration in hours")
```

# Alpha diversity
```{r}
abund_table <-read.csv("SPE_GBS0-12wk.csv",row.names = 1, na.strings="#NULL!")
#Transpose the data to have sample names on rows
abund_table<-t(abund_table)
abund_table <- subset(abund_table,rowSums(abund_table)>11220)

#Extract the corresponding meta_table for the samples in abund_table
meta_a<-meta_table[rownames(abund_table),]

# combine meta and abund tables first
final.a <- data.frame(meta_a, abund_table)
final.a <- final.a[!(rownames(final.a) %in% c("GBS334","GBS346")),]
final.a <- final.a[!(final.a$ipa_indication %in% c(4,5)),] 

# Calculate species richness
N <- rowSums(final.a[,24:ncol(final.a)])

S <- specnumber(final.a[,24:ncol(final.a)])

S.rar <-rarefy(final.a[,24:ncol(final.a)], min(N))
# Calculate community diversity
D <- diversity(final.a[,24:ncol(final.a)])
```

```{r}
PID <- final.a$PID
ipa <- factor(final.a$ipa_indication)
aage <- final.a$actual_age
parity <- final.a$parity
Srar.rirs <- lmer(S.rar ~ aage * parity + (1 + aage | PID))
print(summary(Srar.rirs))
Srar.rirs <- lmer(S.rar ~ aage*ipa + (1 + aage | PID))
#qqnorm(residuals(Srar.rirs))
#qqline(residuals(Srar.rirs))
plot(Srar.rirs,sqrt(abs(resid(.)))~fitted(.),type=c("p","smooth")) # there are outliers here that I would like to exclude
print(summary(Srar.rirs))
aeff <- allEffects(Srar.rirs)
plot(aeff)

#gp <- ggplot(data=final.a, aes(x=aage, y=S.rar, colour=factor(ipa))) + geom_point() + stat_smooth(method = "glm") + theme(axis.title.x = element_blank()) + theme(axis.title.y = element_blank()) + scale_color_manual(values = c("#a6611a","#41b6c4","#225ea8"))
#pdf(file="/home/jstearns/Documents/Baby_mi/richness.pdf", height = 3, width = 5)
#print(gp)
#dev.off()

D.rirs <- lmer(D ~ aage * parity +(1 + aage | PID))
#qqnorm(residuals(D.rirs))
#qqline(residuals(D.rirs))
plot(D.rirs,sqrt(abs(resid(.)))~fitted(.),type=c("p","smooth"))
print(summary(D.rirs))
D.rirs <- lmer(D ~ aage*ipa + aage * parity +(1 + aage | PID))
print(summary(D.rirs))
aeff <- allEffects(D.rirs)
plot(aeff)
    
#gp <- ggplot(data=final.a, aes(x=aage, y=D, colour=factor(ipa))) + geom_point() + stat_smooth(method = "glm") + theme(axis.title.x = element_blank()) + theme(axis.title.y = element_blank()) + scale_color_manual(values = c("#a6611a","#41b6c4","#225ea8"))
#pdf(file="/home/jstearns/Documents/Baby_mi/diversity.pdf", height = 3, width = 5)
#print(gp)
#dev.off()
```


```{r}
# subset samples to compare
# filter OTUs by abundance
# compare abundances with KW otu at a time 
# loop over each OTU
# correct p-values for multiple testing with FDR adjustment
wilcox.test()
```

# Beta diversity
```{r}
L6 <- read.csv("SPE_GBS_L6_nRoot.csv",row.names=1,check.names=FALSE)

L6 <- subset(L6,rowSums(L6) > 0.01)
L6 <- L6[order(rowSums(L6),decreasing = TRUE),]

#Transpose the data to have sample names on rows
L6 <-t(L6)

meta.b <- meta_table[rownames(L6),]

# combine meta
final.b <- data.frame(meta.b, L6)
```

## Vaginal births only

```{r}
vag <- subset(final.b, birth_del_mode==1)
vag <- vag[!(vag$ipa_indication %in% 4),]
vag$birth_del_mode <- factor(vag$birth_del_mode)
vag$ipa_indication <- factor(vag$ipa_indication)

#stratify by age
d3 <- subset(vag, age=="3d")
d10 <- subset(vag, age=="10d")
w6 <- subset(vag, age=="6w")
w12 <- subset(vag, age=="12w")
```

### Multivariate model for permanova
```{r}
ad <- adonis(d3[,24:ncol(d3)] ~ d3$ipa_indication, data=d3[,1:23], method = "bray")
ad

d10.a <- d10[complete.cases(d10[,7]),]
ad <- adonis(d10.a[,24:ncol(d10.a)] ~ d10.a$ipa_indication + d10.a$parity, data=d10.a[,1:23], method = "bray")
ad

w6.a <- w6[complete.cases(w6[,7]),]
ad <- adonis(w6.a[,24:ncol(w6.a)] ~ w6.a$ipa_indication + w6.a$parity, data=w6.a[,1:23], method = "bray")
ad

w12.a <- w12[complete.cases(w12[,7]),]
ad <- adonis(w12.a[,24:ncol(w12.a)] ~ w12.a$ipa_indication + w12.a$parity + w12$birth_del_place, data=w12.a[,1:23], method = "bray")
ad
```

### IPA duration in GBS group alone
```{r}
GBS <- vag[(vag$ipa_indication %in% 3),]

d3.gbs <- GBS[(GBS$age %in% "3d"),]
d10.gbs <- GBS[(GBS$age %in% "10d"),]
w6.gbs <- GBS[(GBS$age %in% "6w"),]
w12.gbs <- GBS[(GBS$age %in% "12w"),]

coefs <- function(ad, thresh){
  df <- t(ad$coefficients)
  #colnames(df) <- c("(Intercept)",var)
  df <- df[order(abs(df[,2]),decreasing = TRUE),]
  #write.csv(df, file = paste(age,var,"coeff.csv",sep = ""))
  df <- df[abs(df[,2]) > thresh,]
}
```

```{r}
# after doing the univariate analysis added everything that was significant into a multivariate model
ad <- adonis(d3.gbs[,24:ncol(d3.gbs)] ~ d3.gbs$ipa_duration, data=d3.gbs[,1:23], method = "bray")
ad
#d3.dur <- coefs(ad,0.005)
#write.csv(d3.dur, "d3.dur.coef.csv")

ad <- adonis(d10.gbs[,24:ncol(d10.gbs)] ~ d10.gbs$ipa_duration, data=d10.gbs[,1:23], method = "bray")
ad
#d10.dur <- coefs(ad,0.005)
#write.csv(d10.dur, "d10.dur.coef.csv")

ad <- adonis(w6.gbs[,24:ncol(w6.gbs)] ~ w6.gbs$ipa_duration + w6.gbs$probiotic_YN, data=w6.gbs[,1:23], method = "bray")
ad
#w6.dur <- coefs(ad,0.005)
#write.csv(w6.dur, "w6.dur.coef.csv")

ad <- adonis(w12.gbs[,24:ncol(w12.gbs)] ~ w12.gbs$ipa_duration + w12.gbs$birth_del_place, data=w12.gbs[,1:23], method = "bray")
ad
#w12.dur <- coefs(ad,0.005)
#write.csv(w12.dur, "w12.dur.coef.csv")
```
at 12 weeks IPA duration accounts for 32% of the variability in the data

## Csection vs vaginal births (noIPA)
```{r}
CSvVG <- final.b[!(final.b$ipa_indication %in% c(3,4,5)),]
CSvVG$ipa_indication <- factor(CSvVG$ipa_indication)

d3.csv <- CSvVG[(CSvVG$age %in% "3d"),]
d10.csv <- CSvVG[(CSvVG$age %in% "10d"),]
w6.csv <- CSvVG[(CSvVG$age %in% "6w"),]
w12.csv <- CSvVG[(CSvVG$age %in% "12w"),]
```

### IPA duration in CS group alone
```{r}
CS <- final.b[(final.b$ipa_indication %in% 2),]
CS$ipa_indication <- factor(CS$ipa_indication)

d3.cs <- CS[(CS$age %in% "3d"),]
d10.cs <- CS[(CS$age %in% "10d"),]
w6.cs <- CS[(CS$age %in% "6w"),]
w12.cs <- CS[(CS$age %in% "12w"),]

coefs <- function(ad, thresh){
  df <- t(ad$coefficients)
  #colnames(df) <- c("(Intercept)",var)
  df <- df[order(abs(df[,2]),decreasing = TRUE),]
  #write.csv(df, file = paste(age,var,"coeff.csv",sep = ""))
  df <- df[abs(df[,2]) > thresh,]
}

d3.cs.a <- d3.cs[complete.cases(d3.cs[,4]),]
ad <- adonis(d3.cs.a[,24:ncol(d3.cs.a)] ~ d3.cs.a$ipa_duration, data=d3.cs.a[,1:23], method = "bray")
ad
#d3.dur <- coefs(ad,0.00008)
#write.csv(d3.dur, "d3.dur.coef.csv")

d10.cs <- d10.cs[complete.cases(d10.cs[,4]),]
ad <- adonis(d10.cs[,24:ncol(d10.cs)] ~ d10.cs$ipa_duration, data=d10.cs[,1:23], method = "bray")
ad
d10.dur <- coefs(ad,0.00008)
#write.csv(d10.dur, "d10cs.dur.coef.csv")
hist(d10.cs$ipa_duration, col = "grey")

w6.cs <- w6.cs[complete.cases(w6.cs[,4]),]
ad <- adonis(w6.cs[,24:ncol(w6.cs)] ~ w6.cs$ipa_duration, data=w6.cs[,1:23], method = "bray")
ad
#w6.dur <- coefs(ad,0.00008)
#write.csv(w6.dur, "w6.dur.coef.csv")
w12.cs <- w12.cs[complete.cases(w12.cs[,4]),]
ad <- adonis(w12.cs[,24:ncol(w12.cs)] ~ w12.cs$ipa_duration, data=w12.cs[,1:23], method = "bray")
ad
#w12.dur <- coefs(ad,0.00008)
#write.csv(w12.dur, "w12.dur.coef.csv")
```
There is a significant effect of ipa duration at day 10 but I'm leary because of the small sample size (n=6).

### Multivariate model for permanova - CS v Vag
```{r}
ad <- adonis(d3.csv[,24:ncol(d3.csv)] ~ d3.csv$birth_del_mode, data=d3.csv[,1:23], method = "bray")
ad

d10.csv.a <- d10.csv[complete.cases(d10.csv[,c(6,7,15)]),]
ad <- adonis(d10.csv.a[,24:ncol(d10.csv.a)] ~ d10.csv.a$birth_del_mode + d10.csv.a$parity + d10.csv.a$probiotic_YN + d10.csv.a$d10_diet_now, data=d10.csv.a[,1:23], method = "bray")
ad

w6.csv.a <- w6.csv[complete.cases(w6.csv[,7]),]
ad <- adonis(w6.csv.a[,24:ncol(w6.csv.a)] ~ w6.csv.a$birth_del_mode + w6.csv.a$parity, data=w6.csv.a[,1:23], method = "bray")
ad

w12.csv.a <- w12.csv[complete.cases(w12.csv[,7]),]
ad <- adonis(w12.csv.a[,24:ncol(w12.csv.a)] ~ w12.csv.a$birth_del_mode + w12.csv.a$parity, data=w12.a[,1:23], method = "bray")
ad
```
At day 3 birth mode is associated with a difference in the microbiome
At day 10 birth mode is associated with a difference in the microbiome when you account for parity, probiotic use and diet
At week 6 birth mode is no longer associated with a difference in the microbiome but parity is (when birth mode is considered).
At week 12 birth mode is associated with a differencein the mirobiome when parity is taken into account.



## Multiple testing correction
```{r}
p <- scan(file = "perm_pvals.txt", what = double()) # this file has a single column of p-values
p.adj <- p.adjust(p, method = "fdr", n = length(p)) # see the documentation for methods
#write(p.adj, "adjusted_pvals.txt", sep = "\n")
pdf <- data.frame(p, p.adjust(p, method = "fdr", n = length(p)))
write.table(pdf,"adjusted_pvals.txt",sep = "\t", row.names = FALSE)
```


# Taxa differences 
```{r}
ipas <- final.b[!(final.b$ipa_indication %in% c(4,5)),]
colnames(ipas[,24:37])
PID <- ipas$PID
actual_age <- ipas$actual_age
ipa <- factor(ipas$ipa_indication)
parity <- factor(ipas$parity)
```

```{r}
gen.rirs <- lmer(ipas$k__Bacteria.p__Actinobacteria.c__Actinobacteria.o__Bifidobacteriales.f__Bifidobacteriaceae.g__Bifidobacterium ~ ipa*actual_age + (1 + actual_age | PID))
print(plot(gen.rirs,sqrt(abs(resid(.)))~fitted(.),type=c("p","smooth")))
print(summary(gen.rirs))
aeff <- allEffects(gen.rirs)
pdf("Bifido.pdf",height=5,width=10)
plot(aeff)
dev.off()

```

```{r}
gen.rirs <- lmer(ipas$k__Bacteria.p__Proteobacteria.c__Gammaproteobacteria.o__Enterobacteriales.f__Enterobacteriaceae.g__Escherichia ~ ipa*actual_age + (1 + actual_age | PID))
print(plot(gen.rirs,sqrt(abs(resid(.)))~fitted(.),type=c("p","smooth")))
print(summary(gen.rirs))
aeff <- allEffects(gen.rirs)
plot(aeff)
pdf("Escherichia.pdf",height=5,width=10)
plot(aeff)
dev.off()
```

```{r}
gen.rirs <- lmer(ipas$k__Bacteria.p__Bacteroidetes.c__Bacteroidia.o__Bacteroidales.f__Bacteroidaceae.g__Bacteroides ~ ipa + actual_age + (1 + actual_age | PID))

print(plot(gen.rirs,sqrt(abs(resid(.)))~fitted(.),type=c("p","smooth")))
print(summary(gen.rirs))
aeff <- allEffects(gen.rirs)
plot(aeff)
pdf("Bacteroides.pdf",height=5,width=10)
plot(aeff)
dev.off()
```

```{r}
gen.rirs <- lmer(ipas$k__Bacteria.p__Firmicutes.c__Bacilli.o__Lactobacillales.f__Streptococcaceae.g__Streptococcus ~ ipa*actual_age + (1 + actual_age | PID))
print(plot(gen.rirs,sqrt(abs(resid(.)))~fitted(.),type=c("p","smooth")))
print(summary(gen.rirs))
aeff <- allEffects(gen.rirs)
plot(aeff)
pdf("Streptococcus.pdf",height=5,width=10)
plot(aeff)
dev.off()
```

```{r}
gen.rirs <- lmer(ipas$k__Bacteria.p__Proteobacteria.c__Gammaproteobacteria.o__Enterobacteriales.f__Enterobacteriaceae.Other ~ ipa*actual_age + (1 + actual_age | PID))
print(plot(gen.rirs,sqrt(abs(resid(.)))~fitted(.),type=c("p","smooth")))
print(summary(gen.rirs))
aeff <- allEffects(gen.rirs)
plot(aeff)
pdf("Enterobacteriaceae.other.pdf",height=5,width=10)
plot(aeff)
dev.off()
```

```{r}
gen.rirs <- lmer(ipas$k__Bacteria.p__Firmicutes.c__Clostridia.o__Clostridiales.f__Lachnospiraceae.g__.Ruminococcus. ~ ipa*actual_age + (1 + actual_age | PID))
print(plot(gen.rirs,sqrt(abs(resid(.)))~fitted(.),type=c("p","smooth")))
print(summary(gen.rirs))
aeff <- allEffects(gen.rirs)
plot(aeff)
pdf("Lachn.Rumino.pdf",height=5,width=10)
plot(aeff)
dev.off()
```

```{r}
gen.rirs <- lmer(ipas$k__Bacteria.p__Firmicutes.c__Clostridia.o__Clostridiales.f__Clostridiaceae.g__Clostridium ~ ipa*actual_age + (1 + actual_age | PID))
print(plot(gen.rirs,sqrt(abs(resid(.)))~fitted(.),type=c("p","smooth")))
print(summary(gen.rirs))
aeff <- allEffects(gen.rirs)
plot(aeff)
pdf("Clostridium.pdf",height=5,width=10)
plot(aeff)
dev.off()
```

```{r}
gen.rirs <- lmer(ipas$k__Bacteria.p__Bacteroidetes.c__Bacteroidia.o__Bacteroidales.f__Porphyromonadaceae.g__Parabacteroides ~ ipa*actual_age + (1 + actual_age | PID))
print(plot(gen.rirs,sqrt(abs(resid(.)))~fitted(.),type=c("p","smooth")))
print(summary(gen.rirs))
aeff <- allEffects(gen.rirs)
plot(aeff)
pdf("Parabacteroides.pdf",height=5,width=10)
plot(aeff)
dev.off()
```

Veillonella (see below)

```{r}
gen.rirs <- lmer(ipas$k__Bacteria.p__Firmicutes.c__Clostridia.o__Clostridiales.f__Clostridiaceae.Other ~ ipa*actual_age + (1 + actual_age | PID))
print(plot(gen.rirs,sqrt(abs(resid(.)))~fitted(.),type=c("p","smooth")))
print(summary(gen.rirs))
aeff <- allEffects(gen.rirs)
plot(aeff)
pdf("Clostridiacea.other.pdf",height=5,width=10)
plot(aeff)
dev.off()
```

```{r}
gen.rirs <- lmer(ipas$k__Bacteria.p__Firmicutes.c__Bacilli.o__Lactobacillales.f__Enterococcaceae.g__Enterococcus ~ ipa*actual_age + (1 + actual_age | PID))
print(plot(gen.rirs,sqrt(abs(resid(.)))~fitted(.),type=c("p","smooth")))
print(summary(gen.rirs))
aeff <- allEffects(gen.rirs)
plot(aeff)
pdf("Enterococcus.pdf",height=5,width=10)
plot(aeff)
dev.off()
```

```{r}
gen.rirs <- lmer(ipas$k__Bacteria.p__Firmicutes.c__Clostridia.o__Clostridiales.f__Lachnospiraceae.Other ~ ipa*actual_age + (1 + actual_age | PID))
print(plot(gen.rirs,sqrt(abs(resid(.)))~fitted(.),type=c("p","smooth")))
print(summary(gen.rirs))
aeff <- allEffects(gen.rirs)
plot(aeff)
pdf("Lachno.pdf",height=5,width=10)
plot(aeff)
dev.off()
```

```{r}
# This is Staphylococcus
gen.rirs <- lmer(ipas$k__Bacteria.p__Firmicutes.c__Bacilli.Other.Other.Other ~ ipa*actual_age + (1 + actual_age | PID))
print(plot(gen.rirs,sqrt(abs(resid(.)))~fitted(.),type=c("p","smooth")))
print(summary(gen.rirs))
aeff <- allEffects(gen.rirs)
plot(aeff)
pdf("Staphylococcus.pdf",height=5,width=10)
plot(aeff)
dev.off()
```

```{r}
gen.rirs <- lmer(ipas$k__Bacteria.p__Proteobacteria.c__Gammaproteobacteria.o__Pasteurellales.f__Pasteurellaceae.g__Haemophilus ~ ipa*actual_age + (1 + actual_age | PID))
print(plot(gen.rirs,sqrt(abs(resid(.)))~fitted(.),type=c("p","smooth")))
print(summary(gen.rirs))
aeff <- allEffects(gen.rirs)
plot(aeff)
```


Based on the Finley paper about the four missing organisms in kids who develop asthma
```{r}
gen.rirs <- lmer(ipas$k__Bacteria.p__Actinobacteria.c__Actinobacteria.o__Actinomycetales.f__Micrococcaceae.g__Rothia ~ ipa*actual_age + (1 + actual_age | PID))
print(plot(gen.rirs,sqrt(abs(resid(.)))~fitted(.),type=c("p","smooth")))
print(summary(gen.rirs))
aeff <- allEffects(gen.rirs)
plot(aeff)
```

```{r}
gen.rirs <- lmer(ipas$k__Bacteria.p__Firmicutes.c__Clostridia.o__Clostridiales.f__Ruminococcaceae.g__Faecalibacterium ~ ipa*actual_age + (1 + actual_age | PID))
print(plot(gen.rirs,sqrt(abs(resid(.)))~fitted(.),type=c("p","smooth")))
print(summary(gen.rirs))
aeff <- allEffects(gen.rirs)
plot(aeff)
```

```{r}
gen.rirs <- lmer(ipas$k__Bacteria.p__Firmicutes.c__Clostridia.o__Clostridiales.f__Lachnospiraceae.g__Lachnospira ~ ipa*actual_age + (1 + actual_age | PID))
print(plot(gen.rirs,sqrt(abs(resid(.)))~fitted(.),type=c("p","smooth")))
print(summary(gen.rirs))
aeff <- allEffects(gen.rirs)
plot(aeff)
```

```{r}
gen.rirs <- lmer(ipas$k__Bacteria.p__Firmicutes.c__Clostridia.o__Clostridiales.f__Veillonellaceae.g__Veillonella ~ ipa*actual_age + (1 + actual_age | PID))
print(plot(gen.rirs,sqrt(abs(resid(.)))~fitted(.),type=c("p","smooth")))
print(summary(gen.rirs))
aeff <- allEffects(gen.rirs)
plot(aeff)
pdf("Veillonella.pdf",height=5,width=10)
plot(aeff)
dev.off()
```