GBS 0-12 weeks: Taxa changes with age. Figure 1
========================================================
Jennifer C. Stearns,
`r as.character(format(Sys.Date(), format="%B %d, %Y"))`

```{r, warning=FALSE, message=FALSE}
setwd("/home/jstearns/Documents/Baby_mi/")
library(knitr)
opts_chunk$set(cache=TRUE)
library(ggplot2)
library(grid)
```

```{r ,echo=FALSE}
# Functions
addKey <- function(table, cols) {
  cbind(ID = interaction(table[cols]),
        table)
}
```

From this [R code for ecology tutorial](http://userweb.eng.gla.ac.uk/umer.ijaz/bioinformatics/ecological.html)
```{r}
# Import meta data
meta_table <- read.csv("sub_meta.csv", row.names = 1, na.strings="#NULL!")

cols = c(5:23)
meta_table[,cols] = apply(meta_table[,cols], 2, factor)

meta_table$age = factor(meta_table$age,levels(meta_table$age)[c(3,1,4,2)])
#hist(meta_table$actual_age, breaks=50,xlim=c(0,110), plot=TRUE,col="grey")
meta_table$ipa_duration <- as.numeric(meta_table$ipa_duration)
```


## Taxa bar charts
### At the Genus level
```{r, echo=FALSE}
# modified the otu table by removing the "Consensus Lineage column" and creating a new OTU ID name for each OTU by concatenating the OTU ID, phyla, family, genus and species name into one with _ between them. Removed spaces and "p__" etc... values.

L6 <-read.csv("SPE_L6_GBS.csv",row.names=1,check.names=FALSE)
L6 <- subset(L6,rowSums(L6) > 0.001)

#Transpose the data to have sample names on rows
L6 <-t(L6)

#meta_table<-read.csv("meta.csv", row.names = 1, na.strings="")
#Just a check to ensure that the samples in meta_table are in the same order as in abund_table
meta_g <- meta_table[rownames(L6),]
final.g <- data.frame(meta_g, L6)

# subset by indication
noIPA.g <- final.g[(final.g$ipa_indication == 1),]
CS.g <- final.g[(final.g$ipa_indication == 2),]
GBS.g <- final.g[(final.g$ipa_indication == 3),]
```


```{r }
colours <- c("#94FFB5","#D05D32","#B75DD5","#83D149","#85BDC5","#586F35","#593B6F","#D0B54A","#81D197","#6E3628","#C74B64","#878BCD","#434749","#C8579D","#F0A3FF","#0075DC","#993F00","#4C005C","#2BCE48","#FFCC99","#808080","#003380","#8F7C00","#9DCC00","#C20088","#190714")#,"#FFA405","#FFA8BB","#426600","#FF0010","#5EF1F2","#00998F","#740AFF","#990000","#FFFF00"

#colours <- c("#642EFE","#689AED","#FE9A2E","#FED733","#1F6397","#A7C49C","#0C9439","#4BEAAE","#F7BE81","#7BA7D0","#7D723E","#AEC359","#676F50","#210B61","#3D728B","#455A65","#F4DDA3","#68AD01","#190714")
```

Bifido "#642EFE" #de2d26
Escherichia "#689AED"
Bacteriodes "#FE9A2E"
Streptococcus "#FED733"
Enterobacteriaceae "#1F6397",
Ruminococcus "#A7C49C",
Veillonella "#0C9439"
Clostridium,"#4BEAAE",
Parabacteriodes "#F7BE81",
Enterococcus "#7BA7D0",
Bacilli "#7D723E",
Clostridiaceae "#AEC359",
Lachnospiraceae "#676F50",
Collinsella "#210B61", #fc9272
Klebsiella "#3D728B",
Haemophilus "#455A65",
Lactobacillus "#F4DDA3",
Megasphaera "#68AD01","#190714"
Neisseria"#434749"
Actinomyces "#593B6F" #fee0d2
Epulospiscium "#BAEB27"
Erysipelotrichaceae "#C3AA09"

```{r, fig.height=7, fig.width=12}
taxa.g <- function(taxa, N, file,w){
  taxa$sample <- rownames(taxa)
  taxa <- addKey(taxa, c(1,152))
  row.names(taxa) <- taxa$ID
  x <- taxa[,25:152]
  x <- x[,order(colSums(x),decreasing=TRUE)]
  #Get grouping information: can add multiple columns here if multiple facets are needed
  grouping_info <- data.frame(row.names=rownames(taxa), as.factor(taxa[,3]))
  #e.g. grouping_info <- data.frame(row.names=rownames(taxa), as.factor(taxa[,3]), as.factor(taxa[,5]))
#--BEGIN OR--
  #Extract list of top N Taxa
  taxa_list <- colnames(x)[1:N]
#Generate a new table with everything added to Others
  new_x <- data.frame(x[,colnames(x) %in% taxa_list],Others=rowSums(x[,!colnames(x) %in% taxa_list]))
#--OR--
#Change any individual value < 1.0% to 0
 #new_x <- x
 #new_x[new_x<0.01] <- 0
#Remove any taxa that now sum to 0
 #new_x <- new_x[,colSums(new_x)>0]
#Add an Others column
 #new_x[,"other"] <- 1-rowSums(new_x)
#--END OR--

  df<-NULL
  for (i in 1:dim(new_x)[2]){
    tmp<-data.frame(row.names=NULL,Sample=rownames(new_x),Taxa=rep(colnames(new_x)[i],dim(new_x)[1]),Value=new_x[,i],Type=grouping_info[,1])
    #tmp<-data.frame(row.names=NULL,Sample=rownames(new_x),Taxa=rep(colnames(new_x)[i],dim(new_x)[1]),Value=new_x[,i],Type=grouping_info[,1], Type2=grouping_info[,2]) if multiple facets are needed
    if(i==1){df<-tmp} else {df<-rbind(df,tmp)}
  }
  
  p<-ggplot(df,aes(Sample,Value,fill=Taxa))+geom_bar(stat="identity")+facet_grid(. ~ Type, drop=TRUE,scale="free",space="free_x")
 #p<-ggplot(df,aes(Sample,Value,fill=Taxa))+geom_bar(stat="identity")+facet_grid(. ~ Type + Type2, drop=TRUE,scale="free",space="free_x") for multiple facets
  p<-p+scale_fill_manual(values=colours[1:(N+1)])
  p<-p+theme_bw()+ylab("Proportions")
  p<-p+ scale_y_continuous(expand = c(0,0))+theme(strip.background = element_rect(fill="gray85"))+theme(panel.margin = unit(0.5, "lines"))# + theme(legend.position="none")
  p<-p+theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5))#+ggtitle(paste("Top",N,"genera over time",sep = " "))
  pdf(file,height=6,width=w)
  print(p)
  dev.off() 
}

colours <- c("#de2d26","#689AED","#FE9A2E","#FED733","#1F6397","#A7C49C","#0C9439","#4BEAAE","#F7BE81","#7BA7D0","#7D723E","#AEC359","#676F50","#fc9272","#3D728B","#455A65","#F4DDA3","#68AD01","#190714")
taxa.g(noIPA.g,18, "noIPA_taxa_2.pdf", 25)

colours <- c("#1F6397","#de2d26","#689AED","#4BEAAE","#FED733","#676F50","#AEC359","#0C9439","#A7C49C","#7D723E","#7BA7D0","#F4DDA3","#434749","#3D728B","#fee0d2","#BAEB27","#455A65","#C3AA09","#190714")
taxa.g(CS.g,18, "CS_taxa_2.pdf", 15)

colours <- c("#689AED","#de2d26","#FE9A2E","#1F6397","#FED733","#F7BE81","#4BEAAE","#455A65","#AEC359","#0C9439","#A7C49C","#3D728B","#7BA7D0","#7D723E","#F4DDA3","#190714")
taxa.g(GBS.g,15, "GBS_taxa_2.pdf", 21)
```

## Effect of IPA at the Phylum level
```{r}
## read in day3 L2 tables for ipa_indication since everyone with GBS was given Penicillin and everyone with a C-section was given Cefazolin
ipa.none <- read.csv("ipa_indication_None.csv",sep="\t", row.names=1)
ipa.csec <- read.csv("ipa_indication_Csec.csv",sep="\t", row.names=1)
ipa.gbs <- read.csv("ipa_indication_GBS.csv",sep="\t", row.names=1)
#par(mfrow=c(1,1), mar=c(5, 4, 4, 2) + 0.1)
#boxplot(ipa.gbs, las=2, at =c(1,2,3,4,5,6, 8,9,10,11,12,13, 15,16,17,18,19,20, 22,23,24,25,26,27), par(mar= c(12, 4, 4, 2) + 0.1), main="Ipa indication = NONE, n=52")

#, par(mar=c(1,4,1,4))
# plotted together
# plotted together
pdf("phyla.pdf",height=5,width=6)
par(mfrow=c(3,1), oma = c(0, 0, 4, 0) + 0.1)
p <- boxplot(ipa.none, las=2, at =c(1,2,3,4,5,6, 8,9,10,11,12,13, 15,16,17,18,19,20, 22,23,24,25,26,27), xaxt="n", par(mar=c(1,4,1,4)), col=c("#D8D8D8","#d73027","#fc8d59","#fee090","#4575b4", "#D8D8D8"))
p <- boxplot(ipa.gbs, las=2, at =c(1,2,3,4,5,6, 8,9,10,11,12,13, 15,16,17,18,19,20, 22,23,24,25,26,27) , xaxt="n", par(mar=c(1,4,1,4)), col=c("#D8D8D8","#d73027","#fc8d59","#fee090","#4575b4", "#D8D8D8"))
p <- boxplot(ipa.csec, las = 2, at =c(1,2,3,4,5,6, 8,9,10,11,12,13, 15,16,17,18,19,20, 22,23,24,25,26,27), xaxt="n", par(mar=c(1,4,1,4)),col=c("#D8D8D8","#d73027","#fc8d59","#fee090","#4575b4", "#D8D8D8"))
p<- legend('topleft', names(ipa.csec)[1:6], pch=15, col=c("#D8D8D8","#d73027","#fc8d59","#fee090","#4575b4", "#D8D8D8"), bty='n', cex=1)
print(p)
dev.off()
```
```