Handy tools
========================================================
Jennifer C. Stearns,
`r as.character(format(Sys.Date(), format="%B %d, %Y"))`

A collection of handy tricks in R
Some are things that I figured out after a bit of heartache and some were adapted from the tutorial [R things I wish I'd been told](http://reganmian.net/blog/2014/10/14/starting-data-analysiswrangling-with-r-things-i-wish-id-been-told/)

```{r}
#Convert to relative frequencies
abund_table<-abund_table/rowSums(abund_table)

# when subseting df by the column id with a list of values in dd
df[df$id %in% dd,]

# when subsetting by multiple values in a column
set <- final[(final$Nethnic_infant %in% c(2, 3)), ]
# exclude multiple values in a column
set <- final[!(final$Nethnic_infant %in% c(2, 3)), ]

# changing the data format of multiple columns in a data frame at once
cols = c(3:7,9:20)
meta_table[,cols] = apply(meta_table[,cols], 2, factor)

# get rid of scientific notation

# subsetting by multiple conditions
my.data.frame <- data[(data$V1 > 2) & (data$V2 < 4), ]
```


```{r}
# from
p <- plot_ordination(L5.data.r, L5.wUF, type="samples", color="Phenotype")
p + geom_point(size = 7) + theme(axis.text=element_text(size=20), axis.title=element_text(size=20),plot.title=element_text(size=20), legend.text=element_text(size=15), legend.title=element_text(size=15)) + scale_color_manual(values=c("#a6611a","#80cdc1","#dfc27d","#018571"))

# make this
my_ordination <- function(data, ord, color, shape=NULL, title) {
  (plot_ordination(
    physeq=data, 
    ordination=ord, 
    type="samples", 
    color=color, 
    shape=shape, 
    title=title) +
  geom_point(size = 7) + 
  theme(axis.text=element_text(size=20), 
        axis.title=element_text(size=20),
        plot.title=element_text(size=20),
        legend.text=element_text(size=15),
        legend.title=element_text(size=15))
  )
}

my_ordination(qiimedata.r, wUF, "ipa_indication", "age", "PCoA of weighted UniFrac distances") # this works
```

Could make a function for preliminary figures and those for publication.

Or cycle through the ordination methods
```{r}
ord_types <- c(wUF, uUF, bc, jac)
color_vars <- c("ipa_indication", "age", "birth_mat_place")
titles <- 
for (groupv in group_vars) {
  for (fillv in fill_vars) {
    my_sideplot(db, groupv, fillv) # this cycles but doesn't plot
  }
}
```

Or cycle through the bacterial groups
```{r}
genera <- c("g__Escherichia","g__Bifidobacterium","g__Streptococcus","g__Bacteroides","f__Enterobacteriaceae","g__Ruminococcus") # or make this list automatically with for (e in names(sample_data(qiimedata.r)[g__Escherichia:g__Ruminococcus])) {do something} or whatever (not tested)
for (gen in genera) {
  my_ordination(data, ord, gen, "age", "PCoA of weighted UniFrac distances")
}
```

Or cycle through the columns in a file (like the meta file!)
```{r}
# not tested... have to change names(db) to names(sample_data(qiimedata)) etc.
for (e in names(db)) {
  if (is.character(db[[e]])) { db[[e]] <- as.factor(db[[e]])}
}
```

Other handy things
```{r}
rownames(x) <- NULL
```

Subsampling the otu table (rarefaction)
```{r}
abund_table <- read.csv("SPE_GBS0-12wk.csv",row.names=1,check.names=FALSE)
#Transpose the data to have sample names on rows
abund_table <- t(abund_table)
#Filter out any samples taxas that have zero entries 
abund_table <- subset(abund_table,rowSums(abund_table)!=0)

depth <- abund_table[order(rowSums(abund_table),decreasing=F),]
head(rowSums(depth))

abund_table <- subset(abund_table,rowSums(abund_table)>11230)

# ugh this is hard can't see how to get it to work
row <- c(1:ncol(abund_table))
N_samples <- 100
for (i in row){
  samples <- sample(1:ncol(abund_table), N_samples, rep=TRUE, prob=abund_table[i,])
}

samples <- sample(1:ncol(abund_table), N_samples, rep=TRUE, prob=abund_table[1,])

# Count the number of each entry and store in a list.
for (i in 1:ncol(abund_table)){
    site_sample[[i]] <- sum(samples==i)
}

# OR try, from (http://ww2.coastal.edu/kingw/statistics/R-tutorials/resample.html)

scores = sleep$extra                           # the data
R = 999                                        # the number of replicates
t.values = numeric(R)                          # storage for the results
for (i in 1:R) {
  group1 = sample(scores, size=10, replace=T)
  group2 = sample(scores, size=10, replace=T)
  t.values[i] = t.test(group1,group2)$statistic
}

```

# plot data
```{r}
# interaction plot
gp <- ggplot(data=final, aes(x=aage, y=S.rar, colour=factor(ipa))) + geom_point() + stat_smooth(method = "glm") + theme(axis.title.x = element_blank()) + theme(axis.title.y = element_blank()) + scale_color_manual(values = c("#a6611a","#41b6c4","#225ea8"))
#pdf(file="/home/jstearns/Documents/Baby_mi/richness.pdf", height = 3, width = 5)
print(gp)
#dev.off()
```
[effects plot with confidence intervals](http://www.r-bloggers.com/plotting-regression-curves-with-confidence-intervals-for-lm-glm-and-glmm-in-r/) This page also has links to pages for checking residuals etc...


# model diagnostics
```{r}
# residuals
plot(tax.rirs,sqrt(abs(resid(.))), fitted(.))
plot(ipa.rirs,sqrt(abs(resid(.)))~fitted(.),type=c("p","smooth"))

plot(fitted(D.lm),residuals(D.lm),xlab="Fitted",ylab="Residuals")
abline(h=0)
lines(smooth.spline(fitted(lmm), residuals(lmm))) # this line should line up with the h=0 line

# normality
qqnorm(residuals(x))
qqline(residuals(Srar.rirs))

# testing model assumptions some simulated data
x <- runif(100, 0, 10)
y <- 1 + 2 * x + rnorm(100, 0, 1)
m <- lm(y ~ x)
par(mfrow = c(2, 2))
plot(m)

```

multiple testing correction
```{r}
p <- scan(file = "pvals.txt", what = double()) # this file has a single column of p-values
p.adj <- p.adjust(p, method = "fdr", n = length(p)) # see the documentation for methods
write(p.adj, "adjusted_pvals.txt", sep = "\n")
```

Longitudinal GLM
```{r}
library(lme4)
library(lmerTest)
# set is data frame
# actual age is the age column
# PID is individual 
gen.rirs <- lmer(set$k__Bacteria.p__Proteobacteria.c__Gammaproteobacteria.o__Enterobacteriales.f__Enterobacteriaceae.g__Escherichia ~ actual_age + (1 + actual_age | PID))
#print(plot(gen.rirs,sqrt(abs(resid(.)))~fitted(.),type=c("p","smooth")))
print(summary(gen.rirs))

library(allEffects)
aeff <- allEffects(gen.rirs)
#pdf("Bifido.pdf",height=5,width=10)
plot(aeff)

```

```{r}
install.packages("factoextra")
library("factoextra")
```

